Why do we Eval?\\[2mm]\Large A large-scale study of Eval usage in R

### 2020-12-01 inst/paper/main.tex
minor

\mypara{Intercession.} A common use case for \eval is to be combined
with \c{match.call}. What \c{match.call} does is step up the call
stack and capture the invocation of the currently executing function
and return it as an unevaluated expression. The general pattern here
is that a call to some function \c{f} is transformed to a call to
another function \c{g} with some arguments retained and others
modified. An illustration taken from the \c{vcpart} package shows a
function \c{tvcglm} that is translated to the \c{tvcm} function but
with two modifications to the argument list, argument \c{control}
can't be missing and argument \c{fit} is forced to have value \c{glm}.
The function ends with a call to \c{eval.parent} to ensure that the
rewritten call is evaluted in the same environment the original call
was.
\begin{lstlisting}
 tvcglm <- function(formula, data, family, control=tvcglm_control(), ...) {
   Call <- match.call()
   Call[[1L]] <- as.name("tvcm")
   if (!"control" %in% names(Call)) Call$control<-formals(tvcglm)$control
   if ("fit" %in% names(list(...))) warning("'fit' is ignored. ")
   Call$fit <- "glm"
   return(eval.parent(Call))
 }
\end{lstlisting}

\mypara{Code Generation.} The more traditional use of \eval is to
execute code that was assembled by the programmer into a string. Here
we show the method \c{plot} for class \c{sback} in package
\c{wsbackfit}, simplified for explanatory purposes. The function takes
a long argument list, the names of which are captured in the list
\c{opt}. The string \c{stub} is composed of a subset of the arguments
passed to this function; the variable is used to construct a call to
\c{base::plot} which will draw a plot. \NOTE{This example is a bit
  messed up. What is var? Does the use of ... in the eval grab the
  arguments of this call? If yes does it not end up with xlab twice?
  HELP}
\begin{lstlisting}
 plot.sback <- function(x,...) {
   opt <- names(list(...))
   stub <- paste(
    ifelse("xlab" %in% opt,"",paste(",xlab=\"",var,"\"",sep="")),
    ifelse("main" %in% opt,"",main.aux),
    ifelse("type" %in% opt,"",",type=\"l\""),
    sep = "")
   plot <- paste("plot(x.data,",stub,",...)", sep = "")
   eval(parse(text=plot))
   ...
\end{lstlisting}
The \parse and \eval combination is used by the the \source function
in the base library to load R code from a file in the current
workspace.

@@

\mypara{Convenience.}
Eval is used to make some programming tasks easier (and/or shorter).
For example, in \c{data.table}, it is used to set R options from a
character vector
  \begin{lstlisting}
opts = c("datatable.verbose"="FALSE", ...) for (i in names(opts))
eval(parse(text=paste0("options(",i,"=",opts[i],")")))

  \end{lstlisting}
  %
 instead of manually calling \c{options(datatable.verbose=FALSE)}

>>>>

\mypara{Intercession.} A common use case for \eval is to be combined
with \c{match.call}. What \c{match.call} does is step up the call
stack and capture the invocation of the currently executing function
and return it as an unevaluated expression. The general pattern here
is that a call to some function \c{f} is transformed to a call to
another function \c{g} with some arguments retained and others
modified. An illustration taken from the \c{vcpart} package shows a
function \c{tvcglm} that is translated to the \c{tvcm} function but
with two modifications to the argument list, argument \c{control}
can't be missing and argument \c{fit} is forced to have value \c{glm}.
The function ends with a call to \c{eval.parent} to ensure that the
rewritten call is evaluted in the same environment the original call
was.
\begin{lstlisting}
 tvcglm <- function(formula, data, family, control=tvcglm_control(), ...) {
   Call <- match.call()
   Call[[1L]] <- as.name("tvcm")
   if (!"control" %in% names(Call)) Call$control<-formals(tvcglm)$control
   if ("fit" %in% names(list(...))) warning("'fit' is ignored. ")
   Call$fit <- "glm"
   return(eval.parent(Call))
 }
\end{lstlisting}
This pattern is recognizable by the fact that the expression is a call
and the target environment is that of the parent.

\mypara{Code Generation.} The more traditional use of \eval is to
execute code that was assembled by the programmer into a string. Here
we show the method \c{plot} for class \c{sback} in package
\c{wsbackfit}, simplified for explanatory purposes. The function takes
a long argument list, the names of which are captured in the list
\c{opt}. The string \c{stub} is composed of a subset of the arguments
passed to this function; the variable is used to construct a call to
\c{base::plot} which will draw a plot. \NOTE{This example is a bit
  messed up. What is var? Does the use of ... in the eval grab the
  arguments of this call? If yes does it not end up with xlab twice?
  HELP} The \parse and \eval combination is used by the the \source
function in the base library to load R code from a file in the current
workspace.
\begin{lstlisting}
 plot.sback <- function(x,...) {
   opt <- names(list(...))
   stub <- paste(
    ifelse("xlab" %in% opt,"",paste(",xlab=\"",var,"\"",sep="")),
    ifelse("main" %in% opt,"",main.aux),
    ifelse("type" %in% opt,"",",type=\"l\""),
    sep = "")
   plot <- paste("plot(x.data,",stub,",...)",sep="")
   eval(parse(text=plot))
   ...
\end{lstlisting}
This pattern is recognized by the use of \c{parse} to turn
a string into an expression.

@@

\mypara{Debloating.} \Eval is often used as a means to reduce
boilerplate code; simple and repetitive code can easily be replaced
with judcious use of \eval. For example, the \c{data.table} package
uses \eval to calls the \c{options} function with named arguments
taken from a vector of strings. While the benefits are limited in this
example, it is an attractive tool for programmers.
\begin{lstlisting}
  opts = c("datatable.verbose"="FALSE", # ... many others
  for (i in names(opts))
    eval(parse(text=paste0("options(",i,"=",opts[i],")")))
\end{lstlisting}
This pattern is a special case of code generation, recognizable by the
fact that \eval is executed in a loop.

\mypara{Trivial.} When values are passed to \eval, they are returned
unchanged. They are an example of trivial uses of \eval. Another
trivial use is the empty expression, often found in JavaScript, but
rare in R.
